---
title: <small>An introduction to digital open <br>research in clinical psychology</small>
author: <small>Matthias Weiler</small>
date: <small>29.03.2019</small>
output:
  revealjs::revealjs_presentation:
    theme: black
    center: TRUE
    self_contained: FALSE
    incremental: TRUE
    reveal_plugins: ["chalkboard"]
    reveal_options:
      chalkboard: 
      readonly: FALSE
---

# <small> What did we do last session?</small>

----

We looked at your research  
proposal presentations

----

... reviewed your proposals

----

... and selected the  
final research projects

----

Any questions or anything to add?



# <small>Last session's homework</small>

----

Who of you added your research project  
to  the "Research projects" section  
of the course in the OSF?

----

... registered your research  
project in the OSF? 

----

... connected your GitHub  
repository to your research  
project in the OSF?

----

Any questions regarding  
last session's homework?



# <small>Today:<br>Data collection</small>

----

## <small>What I was planning to do</small>

- Google Trends
- Twitter
- Sentiment analysis
- Social network analysis

----

## <small>What we are actually doing</small>

- Introduction to R
- Google Trends
- Hellenic Statistical Authority



# <small>Before we begin...</small>

----

If you are still having trouble with  
your RStudio installation, please get  
an account at [https://rstudio.cloud](https://rstudio.cloud)  
now and work there today



# <small>Introduction to R</small>

----

Install packages to get new functions.  
This is how you install the swirl package:  

```{r eval=FALSE}
install.packages("swirl")
```

----

Before you can use the package,  
you have to load the package.  
This is how you load the swirl package:  

```{r eval=FALSE}
library("swirl")
```

----

Start swirl and finish the  
"Basic Building Blocks" exercise  
from the "R Programming" course  
to get in touch with R:

```{r eval=FALSE}
swirl()
```

----

If you are interested, you  
can finish some of the  
other exercises at home



# <small>Group competitions</small>

- Please join your team channel
- Try to find solutions for the  
next exercises as a team  
- Let your team captains post the  
answers in the #data_collection  
channel

----

If you get stuck, I will post hints  
in the #data_collection channel

----

Any questions?



# <small>Let's go</small>



# <small>Google Trends</small>

```{r cache=FALSE, echo=FALSE}
knitr::read_chunk('analysis/google_trends.R')
```


- Find a way to access Google Trends
- Define your search parameters
- Evaluate your data
- Save your results

----

## <small>Find a way to access Google Trends</small>

- Find an R package to access Google Trends
- Post the R code to install the package

----

## <small>Find a way to access Google Trends - Solution</small>

```{r install package, eval=FALSE}
```

----

## <small>Define your search parameters</small>

- Search for the keywords "social phobia"  
and "fobia social"
- ... in Chile and UK
- ... in the year 2018
- Post your query as R code

----

## <small>Define your search parameters - Solution</small>

```{r google trends query}
```

----

## <small>Evaluate your data</small>

- Get the data for interest over time
- Post your R code

----

## <small>Evaluate your data - Solution</small>

```{r get interest over time}
```

----

## <small>Save your results</small>

- Save your results as a csv file
- Post your R code

----

## <small>Save your results - Solution</small>

```{r save data}
```



# <small>Hellenic Statistical Authority</small>

```{r cache=FALSE, echo=FALSE}
knitr::read_chunk('analysis/hell_stat.R')
```

- Find data about deaths in Greece
- Download data
- Import data
- Extract relevant data

----

## <small>Find data about deaths in Greece</small>

- Post the URL (= web address) of the xlsx file  
with causes of death from 2000-2016 from the  
Hellenic Statistical Authority website

----

## <small>Find data about deaths in Greece - Solution</small>

```{}
http://www.statistics.gr/documents/20181/995819/Deaths+-+Causes+of+death+%28ICD-9%29+%28+2000+-+2016+%29/13fcec78-6b8a-4d04-bca6-e84d2d127887?version=1.0
```

----

## <small>Download data</small>

- Find a way to download the xlsx file   
using R code. 
- Post your R code.

----

## <small>Download data - Solution</small>

```{r download data, results='hide'}
```

----

## <small>Import data</small>

- Find a way to import the xlsx file  
using R code. 
- Post your R code.

----

## <small>Import data - Solution</small>

```{r import data, message=FALSE}
```

----

## <small>Extract relevant data</small>

- Extract the number of suicides for  
both genders using R code.  
- Post your R code.

----

## <small>Extract relevant data - Solution</small>

```{r extract relevant data}
```



# <small>Preview: next session</small>

----

"Data analysis"



# <small>Homework</small>

----

Create your own data collection script

----

Upload your script to GitHub



# <small>Any questions?</small>



# <small>Thank you for your attention</small>
